### 2020-11-10 Manuscript/internal_revision/analysis_effects_standard_trajectory.Rmd

---
title: "Analysis of effect sizes in standard trajectory"
#csl: biomed-central.csl
indent: yes
output: bookdown::pdf_document2
toc: false
linestretch: 1.5
fontsize: 12pt
header-includes:
       - \usepackage{subfig}
#bibliography: refs_repro.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(bookdown)
library(knitr)
library(pwr)
library(tidyverse)
```

```{r data_effect_sizes_after_replication, echo = FALSE, message = FALSE, results = 'hide'}
load("effect_sizes_with_CI_after_replication.RData")

# filter all effects > 5 in T1 (standard trajectory)
T1_data <-
  final_init_samp_size_10 %>% 
  filter(trajectory == "T1") %>% 
  filter(effect < 5)

mean_N <- 
  T1_data %>% 
  group_by(distribution) %>% 
  summarize(mean_N = mean(rep_sample_size))

ceiling(mean(T1_data$rep_sample_size))

powered_d_50 <-
  pwr.t.test(n = ceiling(mean(T1_data$rep_sample_size)),
             d = NULL, sig.level = .05, power = .5,
             type = "two.sample",
             alternative = "greater")$d

powered_d_80 <-
  pwr.t.test(n = ceiling(mean(T1_data$rep_sample_size)),
             d = NULL, sig.level = .05, power = .8,
             type = "two.sample",
             alternative = "greater")$d
```

```{r plot_effects_powered_at_50, echo = FALSE, message = FALSE, out.width = "80%", fig.align = 'center'}
ggplot(data = T1_data,
       aes(x = distribution, y = effect)) +
  # facet_wrap(~ distribution) +
  geom_violin(position = position_dodge(width = 0.9)) +
  geom_rect(fill = "lightgrey", alpha = .01, 
            aes(xmin = -Inf, xmax = Inf, 
                ymin = powered_d_50, ymax = Inf)) +
  # geom_rect(fill = "lightgrey", alpha = .01, 
  #           aes(xmin = -Inf, xmax = Inf, 
  #               ymin = powered_d_50, ymax = powered_d_80)) +
  labs(x = "Distribution",
       y = expression(paste("Effect sizes (Cohen's ", italic("d"), ") "))) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text.x = element_text(size = 12, colour = "black")) +
  theme(axis.text.y = element_text(size = 12, colour = "black"))
  # geom_quasirandom(dodge.width = 0.9,
  #                  varwidth = FALSE,
  #                  size = .5, alpha = 0.3)
```

```{r plot_effects_powered_at_80, echo = FALSE, message = FALSE, out.width = "80%", fig.align = 'center'}
ggplot(data = T1_data,
       aes(x = distribution, y = effect)) +
  # facet_wrap(~ distribution) +
  geom_violin(position = position_dodge(width = 0.9)) +
  # geom_rect(fill = "lightgrey", alpha = .01, 
  #           aes(xmin = -Inf, xmax = Inf, 
  #               ymin = powered_d_50, ymax = Inf)) +
  geom_rect(fill = "lightgrey", alpha = .01,
            aes(xmin = -Inf, xmax = Inf,
                ymin = powered_d_80, ymax = Inf)) +
  labs(x = "Distribution",
       y = expression(paste("Effect sizes (Cohen's ", italic("d"), ") "))) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text.x = element_text(size = 12, colour = "black")) +
  theme(axis.text.y = element_text(size = 12, colour = "black"))
  # geom_quasirandom(dodge.width = 0.9,
  #                  varwidth = FALSE,
  #                  size = .5, alpha = 0.3)
```