### 2021-02-05 Manuscript/Supplement.Rmd

---
title: 'Supplement Nature Methods'
csl: nature-methods.csl
indent: yes
output: bookdown::pdf_document2
toc: false
linestretch: 1.5
fontsize: 11pt
header-includes:
       - \usepackage{subfig}
bibliography: refs_repro.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(bookdown)
library(knitr)
library(tidyverse)
```

```{r data, include = FALSE, cache = FALSE}
load("all_outcomes_init_samp_size_10.RData")
load("exploratory_data_SESOI_Carneiro_10.RData")
load("exploratory_data_SESOI_Szucs_10.RData")
load("exploratory_data_significance_Carneiro_10.RData")
load("exploratory_data_significance_Szucs_10.RData")
```

```{r ES data, include = FALSE, cache = FALSE}
load("ES_data_Szucs.RData")
load("ES_data_Carneiro.RData")
```

```{r ES distribution, echo = FALSE, results = "hide"}
min(ES_data_Szucs$D)
max(ES_data_Szucs$D)

sum(ES_data_Szucs$D > 10)

median(ES_data_Szucs$D)
mean(ES_data_Szucs$D)

names(ES_data_Carneiro)[1] <- "D"

ES_data_Carneiro$D <- ifelse(ES_data_Carneiro$D < 0, -ES_data_Carneiro$D, -ES_data_Carneiro$D)

min(ES_data_Carneiro$D)
max(ES_data_Carneiro$D)
median(ES_data_Carneiro$D)
mean(ES_data_Carneiro$D)

ES_data_Carneiro <-
  ES_data_Carneiro %>% 
  mutate(distribution = "Pessimistic")

ES_Szucs <-
  ES_data_Szucs %>% 
  select(D) %>% 
  filter(D <= 10) %>% 
  mutate(distribution = "Optimistic")

ES_data <- bind_rows(ES_Szucs, ES_data_Carneiro)
```

```{r histograms, echo = FALSE, fig.cap = 'Empirical effect size distributions. Note that in (a) 45 values $>$ 10 were removed in order to display the distribution.', fig.subcap = c('Optimistic distribution', 'Pessimistic distribution'), out.width = '.45\\linewidth', fig.ncol = 2, fig.align = 'center'}

ES_histrogram_Szucs <-
  ggplot(data = ES_Szucs, aes(x = D)) +
  geom_histogram(binwidth = .2, color = "black", fill = "white", size = 0.3) +
  theme_bw() +
  labs(x = expression(paste("Empirical effect sizes (Cohen's ", italic("d"), ") ")), 
       y = "Frequency") + 
  scale_x_continuous(breaks = seq(0, 10, 2)) +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black"))

ES_histrogram_Szucs

ES_histrogram_Carneiro <-
  ggplot(data = ES_data_Carneiro, aes(x = D)) +
  geom_histogram(binwidth = .2, color = "black", fill = "white", size = 0.3) +
  theme_bw() +
  labs(x = expression(paste("Empirical effect sizes (Cohen's ", italic("d"), ") ")), 
       y = "Frequency") + 
   scale_x_continuous(breaks = seq(-4, 6, 2)) +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black"))

ES_histrogram_Carneiro
```

```{r, hist, echo = FALSE, fig.cap = 'Exploratory effect sizes (left: based on optimistic empirical distribution, right: based on pessimistic empirical distribution). The panels show the effect sizes that were detected using one of the two decision criteria. Orange shaded bars (panels (a) and (b)) indicate those effect sizes that were identified for replication using the conventional significance threshold (p = .05). Blue shaded bars indicate effect sizes that were selected using a SESOI of 0.5 (panels (c) and (d)) or 1.0 (panels (e) and (f)), respectively. Note that in (a), (c), and (e) 24 values $>$ 10 were removed in order to display the distribution.', fig.align = 'center', fig.subcap = c('', ''), out.width = '.50\\linewidth', fig.ncol = 2}

exploratory_data_sig_Szucs_10 <-
  exploratory_data_sig_Szucs_10 %>%
  filter(pval_threshold == 0.05) %>%
  mutate(reverse_effect = ifelse(effect < 0, -effect, -effect)) %>% 
  filter(reverse_effect <= 10)

ES_histrogram_Szucs <-
  ggplot(data = exploratory_data_sig_Szucs_10,
         aes(x = reverse_effect,
             fill = factor(selection_sig))) +
  geom_histogram(bins = 100, color = "black", size = 0.3) +
  theme_bw() +
  labs(x = expression(paste("Exploratory effect sizes (Cohen's ", italic("d"), ") ")),
       y = "Frequency",
       fill = "Selected for \nreplication") +
  scale_fill_manual(breaks = c("0", "1"),
                    labels = c("no",
                               "yes"),
                    values = c("white", "#E69F00")) +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black")) +
  theme(legend.position = "none")

ES_histrogram_Szucs

exploratory_data_sig_Carneiro_10 <-
  exploratory_data_sig_Carneiro_10 %>% 
  filter(pval_threshold == 0.05) %>% 
  mutate(reverse_effect = ifelse(effect < 0, -effect, -effect))

ES_histrogram_Carneiro_selected_sig <-
  ggplot(data = exploratory_data_sig_Carneiro_10, 
       aes(x = reverse_effect, fill = factor(selection_sig))) +
  geom_histogram(bins = 90, color = "black",
                 size = 0.3, alpha = 0.8) +
  labs(#x = "",
       x = expression(paste("Exploratory effect sizes (Cohen's ", italic("d"), ") ")),
       y = "Frequency",
       fill = "Selected for \nreplication") +
  scale_fill_manual(breaks = c("0", "1"),
                    labels = c("no",
                               "yes"),
                    values = c("white", "#E69F00")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black")) +
  theme(legend.position = "none")

ES_histrogram_Carneiro_selected_sig

exploratory_data_Szucs_10_0.5 <-
  exploratory_data_Szucs_10 %>% 
  filter(SESOI == 0.5) %>% 
  mutate(reverse_effect = ifelse(effect < 0, -effect, -effect)) %>% 
  filter(reverse_effect <= 10)

ES_histrogram_Szucs_selected_SESOI_0.5 <-
  ggplot(data = exploratory_data_Szucs_10_0.5, 
       aes(x = reverse_effect, fill = factor(selection_equiv))) +
  geom_histogram(bins = 90, color = "black",
                 size = 0.3, alpha = 0.8) +
  labs(#x = "",
       x = expression(paste("Exploratory effect sizes (Cohen's ", italic("d"), ") ")),
       y = "Frequency",
       fill = "Selected for \nreplication") +
  scale_fill_manual(breaks = c("0", "1"),
                    labels = c("no",
                               "yes"),
                    values = c("white", "#0072B2")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black")) +
  theme(legend.position = "none")

ES_histrogram_Szucs_selected_SESOI_0.5

exploratory_data_Carneiro_10_0.5 <-
  exploratory_data_Carneiro_10 %>% 
  filter(SESOI == 0.5) %>% 
  mutate(reverse_effect = ifelse(effect < 0, -effect, -effect))

ES_histrogram_Carneiro_selected_SESOI_0.5 <-
  ggplot(data = exploratory_data_Carneiro_10_0.5, 
       aes(x = reverse_effect, fill = factor(selection_equiv))) +
  geom_histogram(bins = 90, color = "black",
                 size = 0.3, alpha = 0.8) +
  labs(#x = "",
       x = expression(paste("Exploratory effect sizes (Cohen's ", italic("d"), ") ")),
       y = "Frequency",
       fill = "Selected for \nreplication") +
  scale_fill_manual(breaks = c("0", "1"),
                    labels = c("no",
                               "yes"),
                    values = c("white", "#0072B2")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black")) +
  theme(legend.position = "none")

ES_histrogram_Carneiro_selected_SESOI_0.5

exploratory_data_Szucs_10_1.0 <-
  exploratory_data_Szucs_10 %>% 
  filter(SESOI == "1.0") %>% 
  mutate(reverse_effect = ifelse(effect < 0, -effect, -effect)) %>% 
  filter(reverse_effect <= 10)

ES_histrogram_Szucs_selected_SESOI_1.0 <-
  ggplot(data = exploratory_data_Szucs_10_1.0, 
       aes(x = reverse_effect, fill = factor(selection_equiv))) +
  geom_histogram(bins = 90, color = "black",
                 size = 0.3, alpha = 0.8) +
  labs(#x = "",
       x = expression(paste("Exploratory effect sizes (Cohen's ", italic("d"), ") ")),
       y = "Frequency",
       fill = "Selected for \nreplication") +
  scale_fill_manual(breaks = c("0", "1"),
                    labels = c("no",
                               "yes"),
                    values = c("white", "#0072B2")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black")) +
  theme(legend.position = "none")

ES_histrogram_Szucs_selected_SESOI_1.0

exploratory_data_Carneiro_10_1.0 <-
  exploratory_data_Carneiro_10 %>% 
  filter(SESOI == "1.0") %>% 
  mutate(reverse_effect = ifelse(effect < 0, -effect, -effect))

ES_histrogram_Carneiro_selected_SESOI_1.0 <-
  ggplot(data = exploratory_data_Carneiro_10_1.0, 
       aes(x = reverse_effect, fill = factor(selection_equiv))) +
  geom_histogram(bins = 90, color = "black",
                 size = 0.3, alpha = 0.8) +
  labs(#x = "",
       x = expression(paste("Exploratory effect sizes (Cohen's ", italic("d"), ") ")),
       y = "Frequency",
       fill = "Selected for \nreplication") +
  scale_fill_manual(breaks = c("0", "1"),
                    labels = c("no",
                               "yes"),
                    values = c("white", "#0072B2")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 16)) +
  theme(axis.title.y = element_text(size = 16)) +
  theme(axis.text.x = element_text(size = 15, colour = "black")) +
  theme(axis.text.y = element_text(size = 15, colour = "black")) +
  theme(legend.position = "none")

ES_histrogram_Carneiro_selected_SESOI_1.0

```

# Supplementary text 1: Limitations