### 2021-02-05 Manuscript/internal_revision/supplemetary_plots.Rmd

---
title: 'Supplementary plots'
#csl: biomed-central.csl
indent: yes
output: bookdown::pdf_document2
toc: false
linestretch: 1.5
fontsize: 12pt
header-includes:
       - \usepackage{subfig}
#bibliography: refs_repro.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(bookdown)
library(knitr)
library(pwr)
library(tidyverse)
```

# PPV for all trajectories and SESOI  
Pessimistic scenario (Carneiro distribution)

```{r data_Carneiro, include = FALSE, cache = FALSE}
outcomes <- load("outcomes_Carneiro_all_trajectories.RData")
```

```{r, PPVplot_Carneiro, echo = FALSE, out.width = "100%", fig.align = 'center'}
facet_names <- 
  c("0.3" = "SESOI = 0.3",
    "0.5" = "SESOI = 0.5",
    "0.7" = "SESOI = 0.7",
    "1" = "SESOI = 1.0")

plot_PPV <- 
  ggplot(data = outcomes_10EU,
         aes(x = trajectory, y = PPV_pop_prev)) + 
  geom_point(size = 1.5) + 
  facet_wrap(~ SESOI, nrow = 2, ncol = 2,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "Trajectory", y = "Positive predictive value") +
  scale_x_discrete(labels = c("SESOI within CI \nSESOI",
                              "Significance \nSESOI",
                              "SESOI within CI \nStandard",
                              "Significance \nStandard")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 11)) +
  theme(axis.title.y = element_text(size = 11)) +
  theme(axis.text.x = element_text(size = 7, colour = "black")) +
  theme(axis.text.y = element_text(size = 10, colour = "black")) +
  theme(strip.text.x = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black"))

hlines <- data.frame(pre_study_odds = c(outcomes_10EU$Prevalence[1], outcomes_10EU$Prevalence[5],
                                        outcomes_10EU$Prevalence[9], outcomes_10EU$Prevalence[13]), 
                     # distribution   = c(rep(plot_data$distribution[1], 2), 
                     #                    rep(plot_data$distribution[5], 2)),
                     SESOI          = rep(c("1", "0.7", "0.5", "0.3")))

plot_PPV <- 
  plot_PPV + 
  geom_hline(data = hlines, 
             aes(yintercept = pre_study_odds),
             color = "red", lty = 2, size = .5)

plot_PPV
```

\newpage

Optimistic scenario (Szucs distribution)

```{r data_Szucs, include = FALSE, cache = FALSE}
outcomes <- load("outcomes_Szucs_all_trajectories.RData")
```

```{r, PPVplot_Szucs, echo = FALSE, out.width = "100%", fig.align = 'center'}
facet_names <- 
  c("0.3" = "SESOI = 0.3",
    "0.5" = "SESOI = 0.5",
    "0.7" = "SESOI = 0.7",
    "1" = "SESOI = 1.0")

plot_PPV <- 
  ggplot(data = outcomes_10EU,
         aes(x = trajectory, y = PPV_pop_prev)) + 
  geom_point(size = 1.5) + 
  facet_wrap(~ SESOI, nrow = 2, ncol = 2,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "Trajectory", y = "Positive predictive value") +
  scale_x_discrete(labels = c("SESOI within CI \nSESOI",
                              "Significance \nSESOI",
                              "SESOI within CI \nStandard",
                              "Significance \nStandard")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 11)) +
  theme(axis.title.y = element_text(size = 11)) +
  theme(axis.text.x = element_text(size = 7, colour = "black")) +
  theme(axis.text.y = element_text(size = 10, colour = "black")) +
  theme(strip.text.x = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black"))

hlines <- data.frame(pre_study_odds = c(outcomes_10EU$Prevalence[1], outcomes_10EU$Prevalence[5],
                                        outcomes_10EU$Prevalence[9], outcomes_10EU$Prevalence[13]), 
                     # distribution   = c(rep(plot_data$distribution[1], 2), 
                     #                    rep(plot_data$distribution[5], 2)),
                     SESOI          = rep(c("1", "0.7", "0.5", "0.3")))

plot_PPV <- 
  plot_PPV + 
  geom_hline(data = hlines, 
             aes(yintercept = pre_study_odds),
             color = "red", lty = 2, size = .5)

plot_PPV
```

```{r}
facet_names <- 
  c("equivalence.SESOI" = "SESOI within CI \nSESOI ",
    "significance.SESOI" = "Significance \nSESOI",
    "equivalence.standard" = "SESOI within CI \nStandard",
    "significance.standard" = "Significance \nStandard")

plot_PPV <- 
  ggplot(data = outcomes_10EU,
         aes(x = SESOI, y = PPV_pop_prev)) + 
  geom_point(size = 1.5) + 
  facet_wrap(~ trajectory, nrow = 2, ncol = 2,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "SESOI", y = "Positive predictive value") +
  # scale_x_discrete(labels = c("SESOI within CI \nSESOI",
  #                             "Significance \nSESOI",
  #                             "SESOI within CI \nStandard",
  #                             "Significance \nStandard")) +
  scale_x_continuous(breaks = c(0.3, 0.5, 0.7, 1.0)) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 11)) +
  theme(axis.title.y = element_text(size = 11)) +
  theme(axis.text.x = element_text(size = 7, colour = "black")) +
  theme(axis.text.y = element_text(size = 10, colour = "black")) +
  theme(strip.text.x = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black"))

# hlines <- data.frame(pre_study_odds = c(outcomes_10EU$Prevalence[1], outcomes_10EU$Prevalence[5],
#                                         outcomes_10EU$Prevalence[9], outcomes_10EU$Prevalence[13]), 
#                      # distribution   = c(rep(plot_data$distribution[1], 2), 
#                      #                    rep(plot_data$distribution[5], 2)),
#                      SESOI          = rep(c("1", "0.7", "0.5", "0.3"))) 
# 
# 
# plot_PPV <- 
#   plot_PPV + 
#   geom_hline(data = hlines, 
#              aes(yintercept = pre_study_odds),
#              color = "red", lty = 2, size = .5)

plot_PPV
```

```{r}
facet_names <- 
  c("equivalence.SESOI" = "SESOI within CI \nSESOI ",
    "significance.SESOI" = "Significance \nSESOI",
    "equivalence.standard" = "SESOI within CI \nStandard",
    "significance.standard" = "Significance \nStandard")

plot_FPR <- 
  ggplot(data = outcomes_10EU,
         aes(x = SESOI, y = FPR)) + 
  geom_point(size = 1.5) + 
  facet_wrap(~ trajectory, nrow = 2, ncol = 2,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "SESOI", y = "False positive rate") +
  # scale_x_discrete(labels = c("SESOI within CI \nSESOI",
  #                             "Significance \nSESOI",
  #                             "SESOI within CI \nStandard",
  #                             "Significance \nStandard")) +
  scale_x_continuous(breaks = c(0.3, 0.5, 0.7, 1.0)) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 11)) +
  theme(axis.title.y = element_text(size = 11)) +
  theme(axis.text.x = element_text(size = 7, colour = "black")) +
  theme(axis.text.y = element_text(size = 10, colour = "black")) +
  theme(strip.text.x = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black"))

# hlines <- data.frame(pre_study_odds = c(outcomes_10EU$Prevalence[1], outcomes_10EU$Prevalence[5],
#                                         outcomes_10EU$Prevalence[9], outcomes_10EU$Prevalence[13]), 
#                      # distribution   = c(rep(plot_data$distribution[1], 2), 
#                      #                    rep(plot_data$distribution[5], 2)),
#                      SESOI          = rep(c("1", "0.7", "0.5", "0.3"))) 
# 
# 
# plot_PPV <- 
#   plot_PPV + 
#   geom_hline(data = hlines, 
#              aes(yintercept = pre_study_odds),
#              color = "red", lty = 2, size = .5)

plot_FPR
```